---
title: "Example Usages of SamplingDesignTools"
author: "Ning Yilin"
date: "Created on 2020-05-29. Updated on `r Sys.Date()`"
output: 
  md_document:
    variant: gfm
---

```{r setup, include=FALSE}
set.seed(1234)
library(knitr)
opts_chunk$set(echo = TRUE, collapse = TRUE, cache = TRUE)
print_percent <- function(prop, n_digits = 1) {
  str <- paste0("%.", n_digits, "f%%")
  sprintf(str, prop * 100)
}
```

# 0 Preparation

Install the SamplingDesignTools package from GitHub (package devtools needed):

```{r}
# devtools::install_github("nyilin/SamplingDesignTools")
```

Load packages used:

```{r}
library(SamplingDesignTools)
library(survival)
library(dplyr)
library(Epi) # To draw (non-counter-matched) nested case-control sample
```

# 1 Drawing counter-matched sample [#S1]

```{r}
data(cohort_1)
head(cohort_1)
m_cox_cohort_1 <- coxph(Surv(t, y) ~ age + gender, data = cohort_1)
```

In this example, I am using a simulated cohort with 10,000 subjects, where a
continuous exposure (x ~ N(55, 10^2)) represents their age and a binary
covariate indicates their gender (c = 1 for male, Pr(c=1)=0.5).

The event time t was generated from a Weibull distribution with hazard
function:

log{h(t; x, c)} = log{h_0} + log(1.1) * x + log(2) * c,

where the baseline hazard h_0=10^{-5} and the scale parameter was 1. All
individuals entered the cohort at time t=0 and were followed until the event
of interest or a censoring time (generated from an exponential distribution with
a rate of 0.05) with a maximum follow-up time of t=25. There were 
`r sum(cohort_1$y == 1)` cases (`r print_percent(mean(cohort_1$y))`) in this
cohort.

## Counter-match on binary surrogate

In this analysis, I dichotomised the continuous exposure at 50 years of age, and
drew a 1:1 time-matched sample counter matched on this binary variable that is
considered as a surrogate for the exposure. This sample was analysed using a
weighted conditional logistic model, where log of sampling weights were
specified as offset in the model.

```{r}
cohort_1$age_bin <- as.numeric(cohort_1$age < 50)
table(cohort_1$age_bin)
ncc_cm_bin <- draw_ncc_cm(cohort = cohort_1, y_name = "y", t_name = "t", 
                          match_var_name = "age_bin", 
                          include_var_name = c("age", "gender"), ml = 1)
head(ncc_cm_bin, 10)
table(ncc_cm_bin$age_bin, ncc_cm_bin$y)
m_clogit_bin <- clogit(y ~ age + gender + strata(set) + offset(log(weight)), 
                       data = ncc_cm_bin)
```

## Counter-match on categorical surrogate

In this analysis, I categorised the continuous exposure into 4 categories at 40,
50 and 60 years of age, and drew a 1:3 time-matched sample counter matched on
this categorical variable that is considered as a surrogate for the exposure.
This sample was again analysed using a weighted conditional logistic model.

```{r}
cohort_1$age_quart <- cut(cohort_1$age, breaks = c(-Inf, 40, 50, 60, Inf), 
                          labels = 1:4, include.lowest = TRUE)
table(cohort_1$age_quart)
ncc_cm_quart <- draw_ncc_cm(cohort = cohort_1, y_name = "y", t_name = "t", 
                            match_var_name = "age_quart", 
                            include_var_name = c("age", "gender"), ml = 1)
head(ncc_cm_quart, 20)
table(ncc_cm_quart$age_quart)
m_clogit_quart <- clogit(y ~ age + gender + strata(set) + offset(log(weight)), 
                         data = ncc_cm_quart)
```

## Compare results

```{r}
results_1 <- rbind(summary(m_cox_cohort_1)$coef, 
                   summary(m_clogit_bin)$coef, 
                   summary(m_clogit_quart)$coef)
rownames(results_1) <- NULL
kable(data.frame(
  Data = c("Full cohort", "", "1:1 NCC-CM", "", "1:3 NCC-CM", ""), 
  Variable = rep(c("Age", "Male"), 3), 
  `True HR` = rep(c(1.1, 2), 3),
  `Estimated HR` = results_1[, "exp(coef)"], 
  `SE of log(HR)` = results_1[, "se(coef)"], 
  `p-value` = results_1[, "Pr(>|z|)"], check.names = FALSE
), digits = c(0, 0, 1, 2, 3, 3))
```

**Notes:** 1:1 NCC-CM indicates the analysis of the time-matched sample
counter-matched on a binary surrogate. 1:3 NCC-CM indicates the analysis
of the time-matched sample counter-matched on a surrogate with 4 categories.

# 2 Compute KM-type weights for NCC sample

```{r}
data(cohort_2)
head(cohort_2)
m_cox_cohort_2 <- coxph(Surv(t, y) ~ x * z + age + gender, data = cohort_2)
```

In this example, I use a simulated cohort with 100,000 subjects with 
age~N(55, 10^2), Pr(male=1)=0.5, a binary exposure (x) that depends
on gender (Pr(x = 1)=0.8 for female and 0.5 for male) and a binary effect 
modifier (Pr(z = 1)=0.3). The event time t was generated from a Weibull
distribution with hazard function:

log{h(t;x, age, gender, z)} = log{h_0} + log(1.5) * x + 
log(1.1) * gender + log(4) * z + log(2) * x * z + 
log(1.02) * age,

where the baseline hazard h_0=0.0005, and the scale parameter was 1. All
individuals entered the cohort at time t=0 and were followed until the event
of interest or a censoring time (generated from an exponential distribution with
a rate of 0.05) with a maximum follow-up time of t=25. There were
`r sum(cohort_2$y == 1)` cases (`r print_percent(mean(cohort_2$y))`) in this
cohort.

## 1:5 NCC matched on age and gender

Subjects were divided into 6 age groups: <35, 36-45, 46-55, 56-65, 66-75 and >75.
A 1:5 NCC sample was draw from the cohort, matched on age group and gender. 

```{r}
n_per_case <- 5
ncc_2 <- ccwc(exit = t, fail = y, controls = n_per_case, 
              match = list(age_cat, gender), include = list(x, age, z), 
              data = cohort_2, silent = TRUE)
names(ncc_2)[-(1:4)] <- c("age_cat", "gender", "x", "age", "z")
head(ncc_2, 12)
# Create the sampling and status indicator
sample_stat <- numeric(nrow(cohort_2))
sample_stat[unique(ncc_2$Map[ncc_2$Fail == 0])] <- 1
sample_stat[ncc_2$Map[ncc_2$Fail == 1]] <- 2
table(sample_stat)
ncc_2_nodup <- compute_km_weights(cohort = cohort_2, t_name = "t", y_name = "y",
                                  sample_stat = sample_stat, 
                                  match_var_names = c("age_cat", "gender"), 
                                  n_per_case = n_per_case)
head(ncc_2_nodup, 12)
summary(ncc_2_nodup$km_weight)
m_cox_ncc_2 <- coxph(Surv(t, y) ~ x * z + gender + age, data = ncc_2_nodup,
                     weights = km_weight, robust = TRUE)
m_clogit_ncc_2 <- clogit(Fail ~ x * z + strata(Set), data = ncc_2)
```

## Compare results

```{r}
results_2 <- rbind(summary(m_cox_cohort_2)$coef, 
                   summary(m_clogit_ncc_2)$coef, 
                   summary(m_cox_ncc_2)$coef[, -3])
results_2 <- data.frame(Variable = rownames(results_2), results_2, 
                        check.names = FALSE)
rownames(results_2) <- NULL
kable(data.frame(
  Data = c("Full cohort", rep("", 4), 
           "1:5 NCC (clogit)", rep("", 2), 
           "1:5 NCC (weighted Cox)", rep("", 4)), 
  Variable = results_2$Variable, 
  `True HR` = c(c(1.5, 4, 1.01, 1.01, 2), 
                c(1.5, 4, 2), c(1.5, 4, 1.01, 1.01, 2)),
  `Estimated HR` = results_2[, "exp(coef)"], 
  `SE of log(HR)` = results_2[, "se(coef)"], 
  `p-value` = results_2[, "Pr(>|z|)"], check.names = FALSE
), digits = c(0, 0, 2, 2, 3, 3))
```

**Notes:** 1:5 NCC indicates the weighted analysis of the NCC without dropping
any controls. 1:5 NCC keep 2 controls indicates the weighted analysis of the NCC
data where only the cases and 2 controls from each set are used. Each NCC data 
was first analysed using conditional logistic regression, and then weighted Cox 
model after breaking the matching.

# 3 Compute weights for NCC without full cohort

In the previous section, we assumed the full cohort was available when computing
the KM-type weight for each subject in the NCC. However, in reality this may not
always be the case. When the cohort is not available, the KM-type weights can be
computed for the NCC sample as long the time of event/censoring for each subject
is available, and the number of subjects at risk can be obtained (or
approximated) elsewhere.

To illustrate this scenario, I reuse the simulated cohort from the previous
section:

```{r}
head(cohort_2)
```

Note that `Time` is the event time of each case in a matched set, while `t` is
the actual event/censoring time of each subject. When the full cohort is not
available, it is often difficult to obtain the exact number of subject at risk
at each event time if the cohort is not available. I simulate such situation
by rounding the event/censoring time (`t`) to the next integer, and use this
coarsened time to compute the number at risk:

```{r}
ncc_2$t <- cohort_2$t[ncc_2$Map]
ncc_2$t_yr <- ceiling(ncc_2$t)
cohort_2$t_yr <- ceiling(cohort_2$t)
risk_table_coarse <- compute_risk_table(cohort = cohort_2, t_name = "t_yr", 
                                        y_name = "y", 
                                        match_var_names = c("age_cat", "gender"))
head(risk_table_coarse)
```

Now I use this information to compute the KM-type weights and subsequently fit
the weighted Cox model:

```{r}
ncc_nodup2 <- compute_km_weights(ncc = ncc_2[, -1], 
                                 risk_table_manual = risk_table_coarse, 
                                 t_name = "t", y_name = "Fail", 
                                 t_match_name = "t_yr",
                                 id_name = "Map", 
                                 match_var_names = c("age_cat", "gender"), 
                                 n_per_case = 5)
m_cox_ncc_2_v2 <- coxph(Surv(t, Fail) ~ x * z + age + gender, 
                        data = ncc_nodup2, weights = km_weight, robust = TRUE)
```

Compare with results when the full cohort is available:

```{r}
results_3 <- rbind(summary(m_cox_cohort_2)$coef, 
                   summary(m_cox_ncc_2)$coef[, -3], 
                   summary(m_cox_ncc_2_v2)$coef[, -3])
results_3 <- data.frame(Variable = rownames(results_3), results_3, 
                        check.names = FALSE)
rownames(results_3) <- NULL
kable(data.frame(
  Data = c("Full cohort", rep("", 4), 
           "1:5 NCC (weighted Cox)", rep("", 4),
           "1:5 NCC (weighted Cox, v2)*", rep("", 4)), 
  Variable = results_3$Variable, 
  `True HR` = rep(c(1.5, 4, 1.01, 1.01, 2), 3),
  `Estimated HR` = results_3[, "exp(coef)"], 
  `SE of log(HR)` = results_3[, "se(coef)"], 
  `p-value` = results_3[, "Pr(>|z|)"], check.names = FALSE
), digits = c(0, 0, 2, 2, 3, 3))
```

*: Using the KM-type weights computed from the approximate number at risk
obtained in this section.

In reality, number at risk at each event time may be approximated by, e.g., size
of the relevant sub-population at mid-year. In such case, user may use the 
following function to generate a template for `risk_table_coarse` to fill in:

```{r}
risk_table_coarse <- prepare_risk_table(ncc = ncc_2, t_match_name = "t_yr", 
                                        y_name = "Fail", 
                                        match_var_names = c("gender", "age_cat"), 
                                        csv_file = NULL)
head(risk_table_coarse)
```

This template will be written to a `csv` if specified by `csv_file`, making it 
easier to supply information regarding the cohort that is required for computing 
the KM-type weights.

# 4 Weighted analysis of (more) extreme case-control samples

In this section, I illustrate the weighted analysis of (more) extreme 
case-control (MECC) samples using the conditional approach described in Section 
2.2 of Salim et al (2014), using the simulated cohort described in Section 1 of
this document:

```{r}
head(cohort_1)
```

For illustrative purpose, I draw a matched MECC sample, with 2 controls
frequency-matched to each case on gender. Unlike a NCC study where any subject
who had the event during the entire follow-up is sampled as a case, in this MECC
study I define cases as subjects who had the event within $\tau_0 = 5$ years of
follow-up, and define eligible controls as subjects who did not have the event
in the first $\tau = 15$ years of follow-up. Note that an event in a MECC study
(`y_mecc`) has a different definition as an event in the original cohort study.

```{r}
set.seed(1)
dat_mecc <- draw_mecc(cohort = cohort_1, tau0 = 5, tau = 15,
                      id_name = "id", t_name = "t", delta_name = "y",
                      match_var_names = "gender", n_per_case = 2)
kable(head(dat_mecc))
table(y = dat_mecc$y, y_mecc = dat_mecc$y_mecc)
```

As the conditional analysis of the MECC sample assumes individual matching, the
function `draw_mecc` randomly matches controls to each case to form matched
sets, indicated by the variable `set_id_mecc`. In addition, this function 
provides the estimated baseline survival probabilities (i.e., $\hat{S}(t_i)$ and 
$\hat{S}(\tau)$) that are needed to compute the weighted likelihood.

The weighted approach estimates the HR, where all covariates need to be centred
at the cohort average:

```{r}
dat_mecc$age_c <- dat_mecc$age - mean(cohort_1$age)
# When there is only a single coefficient in, users should provide reasonable 
# lower and upper limits for the estimate:
m_mecc <- analyse_mecc_cond(
  y_name = "y_mecc", x_formula = ~ age_c, set_id_name = "set_id_mecc",
  surv = dat_mecc$surv, surv_tau = dat_mecc$surv_tau, mecc = dat_mecc,
  lower = -1, upper = 1 
)
round(m_mecc$coef_mat[, -1], 3)
coef_cox_cohort_1 <- summary(m_cox_cohort_1)$coef
kable(data.frame(
  Data = c("Full cohort", "1:2 MECC, weighted"), 
  Variable = "Age", 
  `True HR` = 1.1,
  `Estimated HR` = c(coef_cox_cohort_1["age", "exp(coef)"], 
                     m_mecc$coef_mat["age_c", "exp_est"]), 
  `SE of log(HR)` = c(coef_cox_cohort_1["age", "se(coef)"], 
                     m_mecc$coef_mat["age_c", "se"]), 
  `p-value` = c(coef_cox_cohort_1["age", "Pr(>|z|)"], 
                     m_mecc$coef_mat["age_c", "pval"]), 
  check.names = FALSE
), digits = c(0, 0, 1, 2, 3, 3))
```

**References:**

- Salim A, Ma X, Fall K, et al. Analysis of incidence and prognosis from
'extreme' case–control designs. Stat Med 2014; 33: 5388–5398.
- Støer NC, Salim A, Bokenberger K, et al. Is the matched extreme case–control
design more powerful than the nested case–control design?. Stat Methods Med Res
2019; 28(6): 1911-1923.
